---
title: Using SGS ZOOM accounts
date: September 8, 2025
file: "jds: zoom-accounts/individual-group-use.qmd"
fontsize: 11pt
format:
  sgs-report-typst
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  error = FALSE,
  warning = FALSE
)
library(here)
library(tidyverse)
library(janitor)
library(scales)
library(ggrepel)
library(ourmirror)

# recent charges

# Apr 30, 2025 - May 29	$57.00
# May 30, 2025 - Jun 29	$102.60
# Jun 30, 2025 - Jul 29	$102.60
# Jun 30, 2025 - Jul 29	$94.67
# 2nd quarter total: $356.87

zoom_events <- read_rds(here("data", "zoom_events.rds"))

min_date <- min(zoom_events$month, na.rm = TRUE)
max_date <- max(zoom_events$month, na.rm = TRUE)
date_breaks <- seq.Date(from = min_date, to = max_date, by = "3 months")

jan_2025 <- as.Date("2025-01-01")
if (
  jan_2025 >= min_date && jan_2025 <= max_date && !jan_2025 %in% date_breaks
) {
  date_breaks <- sort(c(date_breaks, jan_2025))
}

custom_labels <- ifelse(
  format(date_breaks, "%m") == "01" | date_breaks == min_date,
  paste(format(date_breaks, "%b"), format(date_breaks, "%Y"), sep = "\n"),
  format(date_breaks, "%b")
)
```

## Background

ZOOM has become an everyday feature of life in Shambhala. We are usually only aware of our own individual uses, not how Zoom is used more broadly. In SGS Zoom is used to offer classes, to host small- and large-scale gatherings, and for administrative functions of all kinds. SGS Zoom accounts are used by SGS staff, Shambhala Online, the Shambhala Board, and other groups such as the Code of Conduct officers and the Global Community Council. We have an educational license, which makes Zoom affordable.

Recently, detailed records of each meeting or event held on Zoom since June 2024 have become available. Those records detail the meeting's host, the number of participants, the event date, whether it was recorded, and its duration. All of this data can be summarized by month to get a sense of the level of activity and trends over time.

```{=typst}
#pagebreak()
```
## Daytime for Shambhala

Here are the busy-times for Shambhala:

```{r}
#| out.width: "100%"
#| fig-height: 6

zoom_heatmap_data <- zoom_events |>
  mutate(
    start_time = start_time + hours(3), # Add 3 hours first
    group = ifelse(
      str_detect(host, "Room"),
      "Shambhala Online",
      "Other Zoom Accounts"
    ),
    hour = hour(start_time),
    day_of_week = wday(start_time, label = TRUE, abbr = FALSE)
  ) |>
  count(group, hour, day_of_week)

# Create the heatmap
ggplot(zoom_heatmap_data, aes(x = hour, y = day_of_week, fill = n)) +
  geom_tile(color = "white", linewidth = 0.5) +
  scale_fill_viridis_c(option = "plasma", name = "Number of\nEvents") +
  scale_x_continuous(breaks = seq(0, 24, by = 3), expand = c(0, 0)) +
  labs(
    title = "Shambhala's Zoom Events",
    subtitle = paste(
      "Total number of events between ",
      month(min_date, label = TRUE),
      year(min_date),
      "and",
      month(max_date, label = TRUE),
      year(max_date)
    ),
    x = "Hour of Day, Eastern Time",
    y = NULL
  ) +
  theme_minimal() +
  facet_wrap(~group, ncol = 1) +
  ourmirror::mirror_theme(base_size = 13) +
  theme(
    panel.grid.major.y = element_blank(),
    axis.text.x = element_text(angle = 0, hjust = 0.5)
    # panel.spacing = unit(.75, "cm")
  )

```

```{=typst}
#pagebreak()
```

## Scale of ZOOM use

Regular staff meetings are the core of SGS administration. (In the following graph, the yellow line represents January 1, 2025.)


```{r}
#| out.width: "100%"

zoom_events |>
  filter(
    str_detect(host, "Tara"),
    str_detect(topic, "Staff"),
    participants > 1
  ) |>
  ggplot(aes(month, participants, color = topic)) +
  geom_vline(
    xintercept = month("2025-01-01"),
    color = "yellow",
    alpha = .5,
    linewidth = 2
  ) +
  geom_point(size = 3) +
  geom_line() +
  scale_x_date(
    labels = custom_labels,
    breaks = date_breaks,
    limits = c(min_date, max_date),
    expand = c(0.02, 0.02)
  ) +
  scale_y_continuous(limits = c(0, 20)) +
  labs(
    title = "Participation in SGS Staff meetings",
    subtitle = "Half of 2024 through August of 2025",
    x = NULL,
    y = NULL,
    color = "Host:"
  ) +
  ourmirror::mirror_theme(base_size = 13)

```

To get a sense of the scale of SGS use, look at the length of zoom meetings (in hours) multiplied by the number of participants. The data doesn't distinguish between internal use within SGS and the hours that our community as a whole spends on ZOOM in events hosted by SGS.

```{r}
#| out.width: 100%

# out.height: 3in
zoom_activity <- zoom_events |>
  filter(participants > 1) |>
  mutate(
    total_hours = (duration_hh_mm_ss * participants) / 3600,
    host = ifelse(
      str_detect(host, "Room") & !str_detect(host, "201"),
      "Shambhla Online",
      "Other"
    )
  ) |>
  summarise(total_hours = sum(total_hours), .by = c(month, host))

zoom_activity |>
  ggplot(aes(month, total_hours, color = host)) +
  geom_vline(
    xintercept = month("2025-01-01"),
    color = "yellow",
    alpha = .5,
    linewidth = 2
  ) +
  geom_point(size = 3) +
  geom_line() +
  scale_x_date(
    labels = custom_labels,
    breaks = date_breaks,
    limits = c(min_date, max_date),
    expand = c(0.02, 0.02)
  ) +
  scale_y_continuous(limits = c(0, 10000), labels = label_comma()) +
  labs(
    title = "Monthly Hours on ZOOM",
    subtitle = "Half of 2024 through August of 2025",
    color = "Host",
    x = NULL,
    y = NULL
  ) +
  ourmirror::mirror_theme(base_size = 13)

```

Some ways of visualizing 10,000 hours in **one month's** use of ZOOM could be:

-   Over 13 months of continuous Zoom video conference calls
-   The energy consumption equivalent to driving an average car approximately 25,000-50,000 miles (2 to years average in the US).

Total video activity in Shambhala is greater than this, since many Zoom recordings are shared after the fact via Vimeo or other storage locations such as Google Drive.

## Recordings

In addition to all the other functions that Zoom offers, recording a meeting to ZOOM's cloud storage is convenient and enables a meeting or class to be reviewed, sharing it with others who missed it live, or for transfer and distribution via Vimeo, in the case of Shambhala Online.

Only 10% of the total zoom meetings between June 2024 and the present were recorded. We pay extra for Zoom recordings that we leave in the cloud for too long. In the 2nd quarter of 2025, SGS paid about \$350 for cloud storage overages.

## Scale Differences in the Use of Recordings

Shambhala Online operates at a much different scale than the rest of SGS in its use of Zoom and the functions of recordings. Shambhala Online has 9 different accounts that are used together and somewhat interchangeably. To show this difference in scale, the following graphs **add up** Shambhala Online activity (where different accounts share an overall function) and contrast it with the individual **averages** of other SGS Zoom accounts (which have distinct functions and are generally not instructional).

```{r}
recordings <- zoom_events |>
  filter(recording == "Yes", minutes(duration_hh_mm_ss) >= 25) |>
  mutate(
    yr_mon = paste(year(month), "-", month(month)),
    host = if_else(str_detect(host, "Room"), "Shambhala Online", host)
  ) |>
  summarise(
    events = n(),
    average_duration = as.numeric(sum(duration_hh_mm_ss) / 3600) / n(),
    total_participants = sum(participants),
    .by = c(yr_mon, month, host)
  )

so_recordings_summary <- recordings |>
  filter(host == "Shambhala Online")

non_so_recordings_summary <- recordings |>
  filter(host != "Shambhala Online") |>
  summarise(
    events = mean(events),
    average_duration = mean(average_duration),
    total_participants = mean(total_participants),
    .by = c(yr_mon, month)
  ) |>
  mutate(host = "All others")

recordings_summary <- bind_rows(
  so_recordings_summary,
  non_so_recordings_summary
)
```

```{=typst}
#pagebreak()
```
In these two graphs, Shambhala Online **totals** are contrasted with **averages** for individual accounts.

```{r}
#| out.width: "100%"
recordings_summary |>
  ggplot(aes(month, events, color = host)) +
  geom_vline(
    xintercept = date("2025-01-01"),
    color = "yellow",
    alpha = .5,
    linewidth = 2
  ) +
  geom_point(size = 3) +
  geom_line() +
  scale_x_date(
    labels = custom_labels,
    breaks = date_breaks,
    limits = c(min_date, max_date),
    expand = c(0.02, 0.02)
  ) +
  scale_y_continuous(limits = c(0, 35)) +
  labs(
    title = "Number of Recorded Events",
    subtitle = "Half of 2024 through August of 2025",
    # caption = "Shambhala Online totals and average for other accounts",
    x = NULL,
    y = NULL,
    color = "Host:"
  ) +
  ourmirror::mirror_theme(base_size = 13)

```

Not only are there **more** Shambhala Online events, those events have more participants.

```{r}
#| out.width: "100%"
recordings_summary |>
  ggplot(aes(month, total_participants, color = host)) +
  geom_vline(
    xintercept = date("2025-01-01"),
    color = "yellow",
    alpha = .5,
    linewidth = 2
  ) +
  geom_point(size = 3) +
  geom_line() +
  scale_x_date(
    labels = custom_labels,
    breaks = date_breaks,
    limits = c(min_date, max_date),
    expand = c(0.02, 0.02)
  ) +
  scale_y_continuous(limits = c(0, 2000), labels = label_comma()) +
  labs(
    title = "Participants in Recorded Events",
    # caption = "Shambhala Online totals and average for other accounts",
    subtitle = "Half of 2024 through August of 2025",
    x = NULL,
    y = NULL,
    color = "Host:"
  ) +
  ourmirror::mirror_theme(base_size = 13)

```

## Participation in Recorded Regularly Scheduled Online Offerings

Regularly-scheduled gatherings presented by Shambhala Online and SGS deparments let us see activity trends.

```{r}
#| out.width: "100%"
#|
individual <- zoom_events |>
  filter(str_detect(topic, "Sunday Gatherings"), hour(duration_hh_mm_ss) >= 1)

individual |>
  filter(participants < 55) |>
  ggplot(aes(month, participants)) +
  geom_vline(
    xintercept = date("2025-01-01"),
    color = "yellow",
    alpha = .5,
    linewidth = 2
  ) +
  geom_smooth(method = "lm", se = FALSE, color = "grey80") +
  geom_point(size = 3) +
  geom_point(
    data = individual |> filter(participants > 55),
    aes(month, participants),
    size = 3
  ) +
  scale_x_date(
    labels = custom_labels,
    breaks = date_breaks,
    limits = c(min_date, max_date),
    expand = c(0.02, 0.02)
  ) +
  scale_y_continuous(limits = c(0, 100)) +
  labs(
    title = "Participation in Sunday Gatherings",
    subtitle = "Half of 2024 through August of 2025",
    caption = "Mamo chants drew many people in February 2024",
    x = NULL,
    y = NULL
  ) +
  ourmirror::mirror_theme(base_size = 13)
```

```{r}
#| out.width: "100%"
individual <- zoom_events |>
  filter(
    str_detect(host, "Tara"),
    str_detect(topic, "Forum"),
    hour(duration_hh_mm_ss) >= 1
  )

individual |>
  ggplot(aes(month, participants)) +
  geom_vline(
    xintercept = date("2025-01-01"),
    color = "yellow",
    alpha = .5,
    linewidth = 2
  ) +
  geom_smooth(method = "lm", se = FALSE, color = "grey80") +
  geom_point(size = 3) +
  # geom_text_repel(aes(label = meeting_number)) +
  scale_x_date(
    labels = custom_labels,
    breaks = date_breaks,
    limits = c(min_date, max_date),
    expand = c(0.02, 0.02)
  ) +
  scale_y_continuous(limits = c(0, 125)) +
  labs(
    title = "Leadership Forums",
    subtitle = "Half of 2024 through August of 2025",
    x = NULL,
    y = NULL
  ) +
  ourmirror::mirror_theme(base_size = 13)
```

```{r}
#| out.width: "100%"
individual <- zoom_events |>
  filter(
    str_detect(host, "Jeff"),
    str_detect(topic, "Leader"),
    participants > 1
  ) |>
  mutate(month = floor_date(month, "month")) |>
  summarise(
    participants = sum(participants),
    .by = month
  )

individual |>
  ggplot(aes(month, participants)) +
  geom_vline(
    xintercept = date("2025-01-01"),
    color = "yellow",
    alpha = .5,
    linewidth = 2
  ) +
  geom_smooth(method = "lm", se = FALSE, color = "grey80") +
  geom_point(size = 3) +
  # geom_text_repel(aes(label = meeting_number)) +
  scale_x_date(
    labels = custom_labels,
    breaks = date_breaks,
    limits = c(min_date, max_date),
    expand = c(0.02, 0.02)
  ) +
  scale_y_continuous(limits = c(0, 100)) +
  labs(
    title = "P&E Leadership Calls",
    subtitle = "Half of 2024 through August of 2025",
    caption = "Three or four P&E calls at different times are offered each quarter.\nHere the number of participants are added up.",
    x = NULL,
    y = NULL
  ) +
  ourmirror::mirror_theme(base_size = 13)
```

There are many other regular events, such as staff meetings, that are core activities but are not regularly recorded.