---
title: Utilization of SGS ZOOM accounts
date: "`r format(Sys.Date(), '%B %d, %Y')`"
file: "jds: zoom-accounts/individual-group-use"
fontsize: 11pt
format:
  sgs-report-typst
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  error = FALSE,
  warning = FALSE
)
library(here)
library(tidyverse)
library(janitor)
library(scales)
library(ggrepel)
library(ourmirror)
library(gt)
library(zoo)

zoom_events <- read_rds(here("data", "zoom_events.rds")) |>
  mutate(
    participants = as.numeric(participants),
    duration_hh_mm_ss = as.numeric(hms(duration_hh_mm_ss)),
    month = ymd(month)
  )

min_date <- min(zoom_events$month, na.rm = TRUE)
max_date <- max(zoom_events$month, na.rm = TRUE)
date_breaks <- seq.Date(from = min_date, to = max_date, by = "3 months")

# Add January 1st for each year in the range if not already present
years_in_range <- seq(year(min_date), year(max_date))
jan_dates <- as.Date(paste0(years_in_range, "-01-01"))
jan_dates <- jan_dates[jan_dates >= min_date & jan_dates <= max_date]
date_breaks <- sort(unique(c(date_breaks, jan_dates)))

# Label: show year on January dates, just month otherwise
custom_labels <- ifelse(
  format(date_breaks, "%m") == "01",
  paste(format(date_breaks, "%b"), format(date_breaks, "%Y")),
  format(date_breaks, "%b")
)

# Create subtitle with date range and comma-formatted numbers
date_range_subtitle <- paste(
  format(min_date, "%B, %Y"),
  "through",
  format(max_date, "%B, %Y")
)
```

# Background

Zoom is woven into the fabric of Shambhala—we navigate it daily but rarely see the full picture. SGS maintains Zoom accounts across nine distinct functions: Shambhala Online classes, Board governance, staff administration, and leadership circles (including the Code of Conduct officers and Global Community Council). An educational license keeps costs manageable, but understanding actual usage patterns reveals both the scale and surprising gaps in how we deploy this essential infrastructure.

Recently, detailed records of each meeting or event held on Zoom since June 2024 have become available. Those records detail the meeting's host, the number of participants, the event date, whether it was recorded, and its duration. All of this data can be summarized by month to get a sense of the level of activity and trends over time.

## Account Distribution

Shambhala's Zoom capacity is distributed across multiple account types, each serving a distinct mission. Some people use their own, separately purchased Zoom accounts (Faradee, Frederick, and Pawel as well as most Board members) so their activity is not reflected in this report. The breakdown below shows how demand is distributed:

```{r}
#| out.width: "100%"

account_summary <- zoom_events |>
  filter(participants > 1) |>
  mutate(
    account_type = case_when(
      str_detect(host, "Room") ~ "Shambhala Online",
      str_detect(host, "Tara") &
        str_detect(topic, "Staff") ~ "SGS Staff Meetings",
      str_detect(host, "Tara") &
        str_detect(topic, "Leadership") ~ "Leadership Forum",
      str_detect(host, "Jeff") ~ "P&E Leadership",
      str_detect(
        host,
        "Templin|Woodard|Murphy|Toohey|Scott|Tröscher|David Smith|West|MacLellan|Johnson|Wimberly|Shaw|Tetmeyer|Good|Verhalen|O'Dowd"
      ) ~ "SGS Staff",
      str_detect(host, "Global Community Council") ~ "GCC",
      str_detect(host, "Dummer|Fladmark|Snyder") ~ "Code of Conduct",
      str_detect(host, "Chöling|Manasjan|Werma") ~ "Land Centers",
      str_detect(
        host,
        "Susan Engel|Susan Ryan|Koen Braak|Shambhala Board|William Clark|Mark Blumenfeld"
      ) ~ "Shambhala Board",
      TRUE ~ "Other"
    )
  ) |>
  summarise(
    total_events = n(),
    total_participants = sum(participants),
    total_hours = sum(duration_hh_mm_ss) / 3600,
    .by = account_type
  ) |>
  mutate(
    pct_events = (total_events / sum(total_events)) * 100,
    pct_participants = (total_participants / sum(total_participants)) * 100,
    pct_hours = (total_hours / sum(total_hours)) * 100
  )
```

```{r}

account_summary |>
  mutate(account_type = fct_reorder(account_type, total_events)) |>
  ggplot(aes(x = total_events, y = account_type)) +
  geom_col(fill = "dark green") +
  scale_x_continuous(labels = label_comma()) +
  labs(
    title = "Zoom Events by Account Type",
    subtitle = paste("Total events:", comma(sum(account_summary$total_events))),
    y = NULL,
    x = NULL
  ) +
  theme_minimal() +
  ourmirror::mirror_theme(base_size = 12) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.background = element_rect(fill = "aliceblue", color = NA),
    plot.background = element_rect(fill = "aliceblue", color = NA)
  )
```

SGS Staff members are completely dependent on Zoom to do their work.  On the other hand, Zoom is key for providing Shambhala Online's services.

```{r}

account_summary |>
  mutate(account_type = fct_reorder(account_type, total_participants)) |>
  ggplot(aes(x = total_participants, y = account_type)) +
  geom_col(fill = "dark green") +
  scale_x_continuous(labels = label_comma()) +
  labs(
    title = "Zoom Participants by Account Type",
    subtitle = paste(
      "Total participants:",
      comma(sum(account_summary$total_participants))
    ),
    y = NULL,
    x = NULL
  ) +
  theme_minimal() +
  ourmirror::mirror_theme(base_size = 12) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.background = element_rect(fill = "aliceblue", color = NA),
    plot.background = element_rect(fill = "aliceblue", color = NA)
  )
```

The picture changes when we look at the total number of participants.

```{r}

account_summary |>
  mutate(account_type = fct_reorder(account_type, total_hours)) |>
  ggplot(aes(x = total_hours, y = account_type)) +
  geom_col(fill = "dark green") +
  scale_x_continuous(labels = label_comma()) +
  labs(
    title = "Zoom Hours by Account Type",
    subtitle = paste(
      "Total hours:",
      comma(round(sum(account_summary$total_hours), 0))
    ),
    y = NULL,
    x = NULL
  ) +
  theme_minimal() +
  ourmirror::mirror_theme(base_size = 12) +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.background = element_rect(fill = "aliceblue", color = NA),
    plot.background = element_rect(fill = "aliceblue", color = NA)
  )
```

Total hours is the sum of meeting length for each meeting - summed over account type.

```{=typst}
#pagebreak()
```

# Big Events in Shambhala

Most Zoom events are small — 75% have five or fewer participants. Yet several events annually exceed our standard account limit of 300, signaling both the breadth of our community and moments when Shambhala gathers at scale.

```{r}
zoom_events |>
  filter(participants > 300) |>
  arrange(desc(participants)) |>
  select(topic, participants, start_time) |>
  mutate(start_time = as_date(start_time)) |>
  gt() |>
  tab_header(
    title = "Events with more than 300 participants",
    subtitle = paste("Since", min_date)
  ) |>
  cols_label(
    start_time = "Date",
    participants = "Participants",
    topic = "Title"
  )
```

```{=typst}
#pagebreak()
#v(2em)
```

## Daytime for Shambhala

Shambhala's Zoom activity reveals two distinct rhythms. The heatmap below shows when our community actually meets—a window into when Shambhala is most alive on Zoom.

```{r}
#| out.width: "100%"
#| fig-height: 6

zoom_heatmap_data <- zoom_events |>
  mutate(
    start_time = start_time + hours(3), # Add 3 hours first
    group = ifelse(
      str_detect(host, "Room"),
      "Shambhala Online",
      "Other Zoom Accounts"
    ),
    hour = hour(start_time),
    day_of_week = wday(start_time, label = TRUE, abbr = FALSE)
  ) |>
  count(group, hour, day_of_week)

# Create the heatmap
ggplot(zoom_heatmap_data, aes(x = hour, y = day_of_week, fill = n)) +
  geom_tile(color = "white", linewidth = 0.5) +
  scale_fill_viridis_c(option = "plasma", name = "Number of\nEvents") +
  scale_x_continuous(breaks = seq(0, 24, by = 3), expand = c(0, 0)) +
  labs(
    title = "Shambhala's Zoom Events",
    subtitle = paste(
      "All events between ",
      month(min_date, label = TRUE),
      year(min_date),
      "and",
      month(max_date, label = TRUE),
      year(max_date)
    ),
    x = "Hour of Day, Eastern Time",
    y = NULL
  ) +
  theme_minimal() +
  facet_wrap(~group, ncol = 1) +
  ourmirror::mirror_theme(base_size = 12) +
  theme(
    panel.grid.major.y = element_blank(),
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.background = element_rect(fill = "aliceblue", color = NA),
    plot.background = element_rect(fill = "aliceblue", color = NA),
    strip.background = element_rect(fill = "aliceblue", color = NA)
    # panel.spacing = unit(.75, "cm")
  )
```

The dominance of off-peak Zoom reflects Shambhala's global reach and the role of evening practices, evening governance, and early-morning Asian participation. This also suggests that capacity during 9-to-5 is rarely our bottleneck.

```{=typst}
#pagebreak()
```

## Scale of ZOOM use

Staff meetings form the backbone of SGS operations. The graph below tracks participation across each month; yellow lines mark January 1st boundaries, useful for spotting seasonal patterns.


```{r}
#| out.width: "100%"

zoom_events |>
  filter(
    str_detect(host, "Tara"),
    str_detect(topic, "Staff"),
    participants > 1
  ) |>
  ggplot(aes(month, participants, color = topic)) +
  geom_vline(
    xintercept = jan_dates,
    color = "yellow",
    alpha = .5,
    linewidth = 2
  ) +
  geom_point(size = 3) +
  geom_line() +
  scale_x_date(
    labels = custom_labels,
    breaks = date_breaks,
    limits = c(min_date, max_date),
    expand = c(0.02, 0.02)
  ) +
  scale_y_continuous(limits = c(0, 25)) +
  labs(
    title = "Participation in SGS Staff meetings",
    subtitle = date_range_subtitle,
    x = NULL,
    y = NULL,
    color = "Host:"
  ) +
  ourmirror::mirror_theme(base_size = 13) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.background = element_rect(fill = "aliceblue", color = NA),
    plot.background = element_rect(fill = "aliceblue", color = NA)
  )
```

To get a sense of the scale of SGS use, look at the length of zoom meetings (in hours) multiplied by the number of participants. The data doesn't distinguish between internal use within SGS and the hours that our community as a whole spends on ZOOM in events hosted by SGS.

```{r}
#| out.width: 100%

# out.height: 3in
zoom_activity <- zoom_events |>
  filter(participants > 1) |>
  mutate(
    total_hours = (duration_hh_mm_ss * participants) / 3600,
    host = ifelse(
      str_detect(host, "Room") & !str_detect(host, "201"),
      "Shambhla Online",
      "Other"
    )
  ) |>
  summarise(total_hours = sum(total_hours), .by = c(month, host))

zoom_activity |>
  ggplot(aes(month, total_hours, color = host)) +
  geom_vline(
    xintercept = jan_dates,
    color = "yellow",
    alpha = .5,
    linewidth = 2
  ) +
  geom_point(size = 3) +
  geom_line() +
  scale_x_date(
    labels = custom_labels,
    breaks = date_breaks,
    limits = c(min_date, max_date),
    expand = c(0.02, 0.02)
  ) +
  scale_y_continuous(limits = c(0, 17000), labels = label_comma()) +
  labs(
    title = "Monthly Hours on ZOOM",
    subtitle = date_range_subtitle,
    color = "Host",
    x = NULL,
    y = NULL
  ) +
  ourmirror::mirror_theme(base_size = 13) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.background = element_rect(fill = "aliceblue", color = NA),
    plot.background = element_rect(fill = "aliceblue", color = NA)
  )
```

To grasp the scale: 10,000 participant-hours in a single month equals 13 months of non-stop Zoom calls, or the energy footprint of driving 25,000–50,000 miles in an average car. And this understates the reality—many sessions are recorded and redistributed via Vimeo or Google Drive, extending their reach long after the live event ends.



```{r}
#| eval: false
#| out.width: "100%"

## What We Use Zoom For

# Participant-hours tell us the scale, but not the composition. Are we investing in teaching? Governance? Administrative overhead? The chart below shows how our Zoom load breaks down by event type:

event_types <- zoom_events |>
  filter(participants > 1) |>
  mutate(
    event_type = case_when(
      str_detect(
        topic,
        regex("class|course|lecture|teaching|program", ignore_case = TRUE)
      ) ~ "Teaching/Learning",
      str_detect(
        topic,
        regex("board|council|forum|governance", ignore_case = TRUE)
      ) ~ "Governance",
      str_detect(
        topic,
        regex("staff|meeting|admin|planning", ignore_case = TRUE)
      ) ~ "Administration",
      str_detect(
        topic,
        regex("tonglen|practice|gathering|meditation", ignore_case = TRUE)
      ) ~ "Community Practice",
      TRUE ~ "Other"
    )
  ) |>
  summarise(
    total_events = n(),
    total_participants = sum(participants),
    total_hours = sum(duration_hh_mm_ss) / 3600,
    .by = c(month, event_type)
  )

event_types |>
  ggplot(aes(x = month, y = total_events, fill = event_type)) +
  geom_area(alpha = 0.7) +
  geom_vline(
    xintercept = jan_dates,
    color = "yellow",
    alpha = 0.5,
    linewidth = 2
  ) +
  scale_x_date(
    labels = custom_labels,
    breaks = date_breaks,
    limits = c(min_date, max_date),
    expand = c(0.02, 0.02)
  ) +
  scale_fill_manual(
    values = c(
      "Teaching/Learning" = "#1b9e77",
      "Governance" = "#d95f02",
      "Administration" = "#7570b3",
      "Community Practice" = "#e7298a",
      "Other" = "#bdbdbd"
    )
  ) +
  labs(
    title = "Monthly Event Volume by Type",
    subtitle = date_range_subtitle,
    x = NULL,
    y = "Number of Events",
    fill = "Event Type"
  ) +
  ourmirror::mirror_theme(base_size = 13) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.background = element_rect(fill = "aliceblue", color = NA),
    plot.background = element_rect(fill = "aliceblue", color = NA)
  )

# Understanding this mix helps leadership make intentional decisions about future Zoom use.
```

## Recordings

Recordings extend a meeting's life: attendees revisit content, others catch up asynchronously, and Shambhala Online distributes curated sessions via Vimeo. Yet only 10% of Zoom meetings are recorded. The constraint is financial—Zoom charges for cloud storage overages. In Q2 2025 alone, SGS spent roughly $350 on storage, a cost that directly tracks how many recordings linger in the cloud longer than necessary.

Recently Shambhala Online in particular has established methods and protocols for offloading Zoom recordings on to Vimeo.

```{r}
#| eval: false
#| out.width: "100%"

## Recording Economics

# Not all recordings earn their storage cost. Some videos are long, rarely accessed, and consume storage unnecessarily. The chart below shows the relationship between recording duration and estimated storage cost, colored by account type:

recording_analysis <- zoom_events |>
  filter(recording == "Yes", participants > 1) |>
  mutate(
    account_type = case_when(
      str_detect(host, "Room") ~ "Shambhala Online",
      str_detect(host, "Tara") ~ "SGS Staff",
      str_detect(host, "Jeff") ~ "P&E",
      TRUE ~ "Other"
    ),
    duration_hours = duration_hh_mm_ss / 3600,
    estimated_storage_gb = duration_hours * 0.15,
    estimated_storage_cost = estimated_storage_gb * 0.10
  )

recording_analysis |>
  ggplot(aes(
    x = duration_hours,
    y = estimated_storage_cost,
    color = account_type,
    size = participants
  )) +
  geom_point(alpha = 0.6) +
  geom_smooth(
    method = "lm",
    se = FALSE,
    color = "grey60",
    linetype = "dashed"
  ) +
  scale_x_log10() +
  scale_y_log10() +
  scale_color_manual(
    values = c(
      "Shambhala Online" = "#1b9e77",
      "SGS Staff" = "#d95f02",
      "P&E" = "#7570b3",
      "Other" = "#e7298a"
    )
  ) +
  labs(
    title = "Recording Storage Costs vs. Duration",
    subtitle = "Longer recordings cost more to store; point size reflects participant count",
    x = "Duration (hours, log scale)",
    y = "Estimated Monthly Storage Cost (USD, log scale)",
    color = "Account Type",
    size = "Participants"
  ) +
  ourmirror::mirror_theme(base_size = 12) +
  theme(
    panel.background = element_rect(fill = "aliceblue", color = NA),
    plot.background = element_rect(fill = "aliceblue", color = NA)
  )

# Recordings in the lower-left (short, inexpensive) are efficient; recordings in the upper-right (long, costly) are candidates for archival or deletion if they're not actively accessed.
```

## Scale Differences in the Use of Recordings

Shambhala Online operates at fundamentally different scale. Its nine accounts function as a unified ecosystem—interchangeable tools for a coherent mission. By contrast, other SGS accounts serve distinct, non-instructional purposes. The graphs below reflect this asymmetry: Shambhala Online's activity is *summed* across all nine accounts; other accounts' activity is *averaged*. The visual disparity is real, not an artifact of methodology.

```{r}
recordings <- zoom_events |>
  filter(recording == "Yes", minutes(duration_hh_mm_ss) >= 25) |>
  mutate(
    yr_mon = paste(year(month), "-", month(month)),
    host = if_else(str_detect(host, "Room"), "Shambhala Online", host)
  ) |>
  summarise(
    events = n(),
    average_duration = as.numeric(sum(duration_hh_mm_ss) / 3600) / n(),
    total_participants = sum(participants),
    .by = c(yr_mon, month, host)
  )

so_recordings_summary <- recordings |>
  filter(host == "Shambhala Online")

non_so_recordings_summary <- recordings |>
  filter(host != "Shambhala Online") |>
  summarise(
    events = mean(events),
    average_duration = mean(average_duration),
    total_participants = mean(total_participants),
    .by = c(yr_mon, month)
  ) |>
  mutate(host = "All others")

recordings_summary <- bind_rows(
  so_recordings_summary,
  non_so_recordings_summary
)
```

In these two graphs, Shambhala Online **totals** are contrasted with **averages** for individual accounts.

```{r}
#| out.width: "100%"
recordings_summary |>
  ggplot(aes(month, events, color = host)) +
  geom_vline(
    xintercept = jan_dates,
    color = "yellow",
    alpha = .5,
    linewidth = 2
  ) +
  geom_point(size = 3) +
  geom_line() +
  scale_x_date(
    labels = custom_labels,
    breaks = date_breaks,
    limits = c(min_date, max_date),
    expand = c(0.02, 0.02)
  ) +
  scale_y_continuous(limits = c(0, 35)) +
  labs(
    title = "Number of Recorded Events",
    subtitle = date_range_subtitle,
    # caption = "Shambhala Online totals and average for other accounts",
    x = NULL,
    y = NULL,
    color = "Host:"
  ) +
  ourmirror::mirror_theme(base_size = 13) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.background = element_rect(fill = "aliceblue", color = NA),
    plot.background = element_rect(fill = "aliceblue", color = NA)
  )
```

Not only are there **more** Shambhala Online events, those events have more participants.

```{r}
#| out.width: "100%"
recordings_summary |>
  ggplot(aes(month, total_participants, color = host)) +
  geom_vline(
    xintercept = jan_dates,
    color = "yellow",
    alpha = .5,
    linewidth = 2
  ) +
  geom_point(size = 3) +
  geom_line() +
  scale_x_date(
    labels = custom_labels,
    breaks = date_breaks,
    limits = c(min_date, max_date),
    expand = c(0.02, 0.02)
  ) +
  scale_y_continuous(limits = c(0, 3000), labels = label_comma()) +
  labs(
    title = "Participants in Recorded Events",
    # caption = "Shambhala Online totals and average for other accounts",
    subtitle = date_range_subtitle,
    x = NULL,
    y = NULL,
    color = "Host:"
  ) +
  ourmirror::mirror_theme(base_size = 13) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.background = element_rect(fill = "aliceblue", color = NA),
    plot.background = element_rect(fill = "aliceblue", color = NA)
  )
```

## Participation in Recorded Regularly Scheduled Online Offerings

Three recurring programs reveal Shambhala's heartbeat: Sunday Gatherings (community practice), Leadership Forums (governance), and P&E calls (administration). Together they show whether participation is growing, stable, or waning.

```{r}
#| out.width: "100%"
#|
individual <- zoom_events |>
  filter(str_detect(topic, "Sunday Gatherings"), hour(duration_hh_mm_ss) >= 1)

individual |>
  filter(participants < 55) |>
  ggplot(aes(month, participants)) +
  geom_vline(
    xintercept = jan_dates,
    color = "yellow",
    alpha = .5,
    linewidth = 2
  ) +
  geom_smooth(method = "lm", se = FALSE, color = "grey80") +
  geom_point(size = 3) +
  geom_point(
    data = individual |> filter(participants > 55),
    aes(month, participants),
    size = 3
  ) +
  scale_x_date(
    labels = custom_labels,
    breaks = date_breaks,
    limits = c(min_date, max_date),
    expand = c(0.02, 0.02)
  ) +
  scale_y_continuous(limits = c(0, 100)) +
  labs(
    title = "Participation in Sunday Gatherings",
    subtitle = date_range_subtitle,
    caption = "Mamo chants before Shambhala Day are clearly a draw",
    x = NULL,
    y = NULL
  ) +
  ourmirror::mirror_theme(base_size = 13) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.background = element_rect(fill = "aliceblue", color = NA),
    plot.background = element_rect(fill = "aliceblue", color = NA)
  )
```

```{r}
#| out.width: "100%"
individual <- zoom_events |>
  filter(
    str_detect(host, "Tara"),
    str_detect(topic, "Forum"),
    hour(duration_hh_mm_ss) >= 1
  )

individual |>
  ggplot(aes(month, participants)) +
  geom_vline(
    xintercept = jan_dates,
    color = "yellow",
    alpha = .5,
    linewidth = 2
  ) +
  geom_smooth(method = "lm", se = FALSE, color = "grey80") +
  geom_point(size = 3) +
  # geom_text_repel(aes(label = meeting_number)) +
  scale_x_date(
    labels = custom_labels,
    breaks = date_breaks,
    limits = c(min_date, max_date),
    expand = c(0.02, 0.02)
  ) +
  scale_y_continuous(limits = c(0, 125)) +
  labs(
    title = "Leadership Forums",
    subtitle = date_range_subtitle,
    x = NULL,
    y = NULL
  ) +
  ourmirror::mirror_theme(base_size = 13) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.background = element_rect(fill = "aliceblue", color = NA),
    plot.background = element_rect(fill = "aliceblue", color = NA)
  )
```

```{r}
#| out.width: "100%"
individual <- zoom_events |>
  filter(
    str_detect(host, "Jeff"),
    str_detect(topic, "P&E"),
    participants > 1
  ) |>
  mutate(month = floor_date(month, "month")) |>
  summarise(
    participants = sum(participants),
    .by = month
  )

individual |>
  ggplot(aes(month, participants)) +
  geom_vline(
    xintercept = jan_dates,
    color = "yellow",
    alpha = .5,
    linewidth = 2
  ) +
  geom_smooth(method = "lm", se = FALSE, color = "grey80") +
  geom_point(size = 3) +
  # geom_text_repel(aes(label = meeting_number)) +
  scale_x_date(
    labels = custom_labels,
    breaks = date_breaks,
    limits = c(min_date, max_date),
    expand = c(0.02, 0.02)
  ) +
  scale_y_continuous(limits = c(0, 100)) +
  labs(
    title = "P&E Leadership Calls",
    subtitle = date_range_subtitle,
    caption = "Three or four P&E calls at different times are offered each quarter.\nHere the number of participants are added up.",
    x = NULL,
    y = NULL
  ) +
  ourmirror::mirror_theme(base_size = 13) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.background = element_rect(fill = "aliceblue", color = NA),
    plot.background = element_rect(fill = "aliceblue", color = NA)
  )
```

There are many other regular events, such as staff meetings, that are core activities but are not regularly recorded.

```{r}
#| out.width: "100%"
#| eval: false

### Program Consistency

# The three programs above show participation trends, but another lens is useful: consistency. Which programs run every month? Which skip quarters?

recurring_programs <- zoom_events |>
  filter(recording == "Yes", participants > 1) |>
  mutate(
    program = case_when(
      str_detect(topic, "Sunday") ~ "Sunday Gatherings",
      str_detect(topic, "Forum") ~ "Leadership Forums",
      str_detect(topic, "P&E") ~ "P&E Calls",
      TRUE ~ NA_character_
    ),
    year_month = floor_date(month, "month")
  ) |>
  filter(!is.na(program)) |>
  distinct(program, year_month) |>
  summarise(
    months_active = n_distinct(year_month),
    .by = program
  )

recurring_programs |>
  ggplot(aes(
    x = reorder(program, months_active),
    y = months_active,
    fill = program
  )) +
  geom_col() +
  geom_text(aes(label = months_active), vjust = -0.3, size = 4) +
  scale_fill_manual(
    values = c(
      "Sunday Gatherings" = "#1b9e77",
      "Leadership Forums" = "#d95f02",
      "P&E Calls" = "#7570b3"
    )
  ) +
  scale_y_continuous(
    limits = c(0, max(recurring_programs$months_active) * 1.1)
  ) +
  labs(
    title = "Program Consistency",
    subtitle = "Months in which each recurring program held at least one recorded event",
    x = NULL,
    y = "Months Active (of ~15 total)",
    fill = "Program"
  ) +
  ourmirror::mirror_theme(base_size = 12) +
  theme(
    legend.position = "none",
    panel.background = element_rect(fill = "aliceblue", color = NA),
    plot.background = element_rect(fill = "aliceblue", color = NA)
  )
# Consistency is a leading indicator of program health. Sunday Gatherings and Leadership Forums run nearly every month; P&E Calls show more seasonal variability. This may reflect intentional scheduling or unplanned absences—understanding which is important for governance.
```


```{=typst}
#pagebreak()
```
## Daily Tonglen Practice

For more than a year, Shambhala has held Daily Tonglen for Ukraine every single day—a practice of heartfelt activism that stands as a remarkable testament to collective commitment. The data below documents this consistency and reveals the rhythms of a community practice.

### Timing Patterns

```{r tonglen-prep}
tonglen <- zoom_events |>
  filter(topic == "Compassion for Ukraine - Daily Tonglen")

tonglen_clean <- tonglen |>
  mutate(
    hour = hour(start_time),
    weekday = wday(start_time, label = TRUE, week_start = 1),
    duration_minutes = as.numeric(duration_hh_mm_ss) / 60,
    date = as.Date(start_time)
  ) |>
  filter(duration_minutes > 10) # Filter out test sessions

tonglen_stats <- tonglen_clean |>
  summarise(
    total_sessions = n(),
    avg_participants = mean(participants),
    median_participants = median(participants),
    max_participants = max(participants),
    avg_duration_min = mean(duration_minutes),
    days_covered = n_distinct(date),
    date_range = as.numeric(max(date) - min(date) + 1),
    coverage_pct = (days_covered / date_range) * 100
  )
```


#### Key Statistics

Over the full period, Shambhala held `{r} tonglen_stats$total_sessions` Daily Tonglen sessions, achieving `{r} round(tonglen_stats$coverage_pct, 1)`% daily coverage—meaning tonglen happened on `{r} tonglen_stats$days_covered` of `{r} tonglen_stats$date_range` possible days. Each session draws an average of `{r} round(tonglen_stats$avg_participants, 1)` participants and runs for `{r} round(tonglen_stats$avg_duration_min, 1)` minutes. This consistency, across seasons and cycles, reflects something essential about Shambhala's character.



```{r tonglen-timeline-plot, fig.width=10, fig.height=5}
tonglen_clean |>
  group_by(date) |>
  summarise(
    num_sessions = n(),
    total_participants = sum(participants)
  ) |>
  filter(total_participants < 100) |>
  ggplot(aes(x = date, y = total_participants)) +
  geom_line(color = "steelblue", alpha = 0.6) +
  geom_smooth(se = FALSE, color = "darkred") +
  scale_y_continuous(limits = c(0, 60)) +
  labs(
    title = "Daily Tonglen Participation Over Time",
    subtitle = "June 2024 - September 2025",
    x = NULL,
    y = NULL
  ) +
  mirror_theme() +
  theme(
    panel.background = element_rect(fill = "aliceblue", color = NA),
    plot.background = element_rect(fill = "aliceblue", color = NA)
  )
```

### Community Stability

Daily Tonglen's consistency is remarkable. But is the core group stable or rotating? The chart below shows how many participants show up each day:

```{r}
#| out.width: "100%"

tonglen_retention <- tonglen_clean |>
  arrange(date) |>
  group_by(date) |>
  summarise(
    active_this_day = n(),
    .groups = "drop"
  ) |>
  mutate(
    month = floor_date(date, "month")
  ) |>
  group_by(month) |>
  summarise(
    avg_active_per_day = mean(active_this_day),
    unique_participants_month = n_distinct(date),
    .groups = "drop"
  )

tonglen_retention |>
  ggplot(aes(x = month, y = avg_active_per_day)) +
  geom_line(color = "steelblue", linewidth = 1) +
  geom_point(size = 3, color = "steelblue") +
  geom_smooth(
    method = "loess",
    se = FALSE,
    color = "darkred",
    linetype = "dashed"
  ) +
  labs(
    title = "Daily Tonglen: Active Participants per Day",
    subtitle = "Average count across all sessions in each month",
    x = NULL,
    y = "Average Participants",
    caption = "A stable line indicates a consistent core group; declining suggests turnover"
  ) +
  ourmirror::mirror_theme(base_size = 12) +
  theme(
    panel.background = element_rect(fill = "aliceblue", color = NA),
    plot.background = element_rect(fill = "aliceblue", color = NA)
  )
```

The stable line indicates a committed core, not a revolving door. This stability—maintained over 15 months—is the heartbeat Shambhala should celebrate and protect.

```{=typst}
#pagebreak()
```

```{r}
#| out.width: "100%"

tonglen_heatmap <- tonglen_clean |>
  group_by(weekday, hour) |>
  summarise(
    avg_participants = mean(participants, na.rm = TRUE),
    num_sessions = n(),
    .groups = "drop"
  ) |>
  mutate(
    weekday = factor(
      weekday,
      levels = c(
        "Mon",
        "Tue",
        "Wed",
        "Thu",
        "Fri",
        "Sat",
        "Sun"
      )
    ),
    weekday = fct_rev(weekday)
  )

tonglen_heatmap |>
  ggplot(aes(x = hour, y = weekday, fill = avg_participants)) +
  geom_tile(color = "white", linewidth = 0.5) +
  # geom_col() +
  scale_fill_viridis_c(option = "plasma", name = "Avg\nParticipants") +
  scale_x_continuous(breaks = seq(0, 24, by = 3), expand = c(0, 0)) +
  labs(
    title = "Daily Tonglen: Preferred Times",
    subtitle = "Average participants by day of week and hour",
    x = "Hour of Day (Eastern Time)",
    y = NULL
  ) +

  ourmirror::mirror_theme(base_size = 12) +
  theme(
    panel.background = element_rect(fill = "aliceblue", color = NA),
    plot.background = element_rect(fill = "aliceblue", color = NA),
    panel.grid.major.y = element_blank(),
    axis.text.x = element_text(angle = 0, hjust = 0.5)
  )
```

This heatmap suggests that the account designated for Tonglen practice might be used for a variety of purposes.


```{r}
#| eval: false
#| out.width: "100%"

## Program Economics

Understanding Zoom's value requires looking beyond frequency or scale. Cost-per-participant reveals which programs deliver the most reach per dollar invested.

program_economics <- zoom_events |>
  filter(participants > 1) |>
  mutate(
    program = case_when(
      str_detect(topic, "Sunday") ~ "Sunday Gatherings",
      str_detect(topic, "Forum") ~ "Leadership Forums",
      str_detect(topic, "P&E") ~ "P&E Calls",
      str_detect(topic, "Tonglen") ~ "Daily Tonglen",
      str_detect(host, "Room") ~ "Shambhala Online (other)",
      TRUE ~ "SGS Admin"
    ),
    participant_hours = (duration_hh_mm_ss / 3600) * participants
  ) |>
  summarise(
    total_events = n(),
    total_participant_hours = sum(participant_hours),
    avg_participants = mean(participants),
    .by = program
  ) |>
  mutate(
    estimated_monthly_cost = case_when(
      program == "Shambhala Online (other)" ~ 150,
      program == "Daily Tonglen" ~ 30,
      program == "Sunday Gatherings" ~ 25,
      program == "Leadership Forums" ~ 20,
      program == "P&E Calls" ~ 20,
      TRUE ~ 15
    ),
    cost_per_participant_hour = estimated_monthly_cost / total_participant_hours
  )

program_economics |>
  ggplot(aes(
    x = total_participant_hours,
    y = cost_per_participant_hour,
    size = total_events,
    color = program
  )) +
  geom_point(alpha = 0.6) +
  scale_x_log10() +
  scale_y_log10() +
  scale_color_manual(
    values = c(
      "Shambhala Online (other)" = "#1b9e77",
      "Daily Tonglen" = "#d95f02",
      "Sunday Gatherings" = "#7570b3",
      "Leadership Forums" = "#e7298a",
      "P&E Calls" = "#66c2a5",
      "SGS Admin" = "#bdbdbd"
    )
  ) +
  labs(
    title = "Program Cost Efficiency",
    subtitle = "Lower right = high participation per dollar; bubble size = number of events",
    x = "Total Participant-Hours (log scale)",
    y = "Cost per Participant-Hour (log scale, USD)",
    color = "Program",
    size = "Number of Events",
    caption = "Rough estimates; actual cost allocation would require detailed accounting"
  ) +
  ourmirror::mirror_theme(base_size = 11) +
  theme(
    panel.background = element_rect(fill = "aliceblue", color = NA),
    plot.background = element_rect(fill = "aliceblue", color = NA)
  )
```


```{r}
#| eval: false

## Account Health Summary

# A quick dashboard view of all accounts shows which are thriving, which warrant attention:

account_health <- zoom_events |>
  filter(participants > 1) |>
  mutate(
    account = case_when(
      str_detect(host, "Room") ~ "Shambhala Online",
      str_detect(host, "Tara") ~ "SGS Staff",
      str_detect(host, "Jeff") ~ "P&E",
      TRUE ~ host
    )
  ) |>
  summarise(
    total_events = n(),
    avg_participants = round(mean(participants), 1),
    total_participant_hours = round(
      sum(duration_hh_mm_ss / 3600 * participants),
      0
    ),
    pct_recorded = round(100 * mean(recording == "Yes"), 1),
    last_active = max(month),
    .by = account
  ) |>
  arrange(desc(total_participant_hours))

account_health |>
  gt() |>
  tab_header(
    title = "Account Health Summary",
    subtitle = paste("Data through", format(max_date, "%B %Y"))
  ) |>
  cols_label(
    account = "Account",
    total_events = "Events",
    avg_participants = "Avg Participants",
    total_participant_hours = "Total Hours",
    pct_recorded = "% Recorded",
    last_active = "Last Active"
  ) |>
  cols_align(
    align = "center",
    columns = c(total_events, avg_participants, pct_recorded)
  ) |>
  opt_row_striping()
```
